%---------------------------------------------------------------------------
% Preface

\chapter*{Preface}

The work presented in this thesis was conducted within the Information Science and Engineering 
group at the Institute for Machine Learning (ETH Zurich), under the supervision of 
Prof. Dr. Joachim M. Buhmann. The thesis was co-supervised at Universitat Polit\`ecnica 
de Catalunya by Prof. Dr. Alexandre Parera i Lluna.

\chapter*{Acknowledgements}

First and foremost, I would like to express my deepest gratitude to Prof. Joachim M. 
Buhmann for granting me the opportunity to join the ISE group. I must specifically 
acknowledge his original work on Posterior Agreement, which constitutes the cornerstone 
of the ideas and methods explored in this thesis. The first two chapters rely heavily 
on his foundational contributions, and it is impossible to fully do justice to the 
breadth of his work within the text. I would also like to thank him for his guidance, 
as well as for the many interesting insights and conversations we shared. As he embarks on his 
well-deserved retirement, I wish him all the best for the future. \\

I am equally grateful to Dr. Jo\~ao Carvalho, whose mentorship has greatly contributed 
to my growth, both academically and personally. I sincerely appreciate the opportunity 
he gave me to join this project, as well as for involving me in other exciting and 
interesting endeavors. His valuable insights have had a profound impact on the 
work presented here and beyond. I am truly thankful for everything he has provided, 
and I will carry the lessons learned from him throughout my life. \\

My sincere thanks also go to Dr. Alessandro Torcinovich, whose advice and experience have consistently 
helped refine and improve my research skills. His thoughtful feedback has greatly enhanced 
the quality of this thesis and my future work as well. I truly appreciate the time he 
dedicated to guiding me and offering insightful ideas that helped further 
develop this project. \\

Finally, I would like to extend my heartfelt thanks to all the members of the ISE group,
namely Eugene, Fabian, Ivan, Robin, Xia, Lukas, Ami, Alina and Rita. Your warm acceptance 
and collaboration made my time here not only productive but truly enjoyable. 
In one way or another, each of you helped create a stimulating research environment, 
and I am grateful for the many engaging conversations we shared. I leave with 
cherished memories of our time together and wish you all continued success in your 
future endeavors. \\
 
 \cleardoublepage

%---------------------------------------------------------------------------
% Table of contents

 \setcounter{tocdepth}{2}
 \tableofcontents

 \cleardoublepage

%---------------------------------------------------------------------------
% Abstract

\chapter*{Abstract}
 \addcontentsline{toc}{chapter}{Abstract}

 Posterior Agreement (PA) has been proposed as a theoretically-grounded alternative 
 for model robustness assessment in covariate shift settings. In this work, we provide 
 further evidence in favor of these findings, and we explore the use of PA as a model 
 selection criterion for deep learning models in supervised image classification tasks. \\
 
 Starting from the theoretical principles leading to the formulation of PA, we derive an
 operative version for discrete hypothesis set problems and use it as a robustness metric 
 in the presence of adversarially-perturbed and domain-shifted data. We show that PA outperforms 
 standard accuracy-based metrics in both settings and possesses superior discriminative 
 power and consistency across increasing levels of covariate shift. \\

 \cleardoublepage


%---------------------------------------------------------------------------
% Symbols

% \chapter*{Notation}\label{chap:symbole}
%  \addcontentsline{toc}{chapter}{Notation}

% \section*{Symbols}
% \begin{tabbing}
%  \hspace*{1.6cm} \= \hspace*{8cm} \= \kill
%  $\mathrm{EHC}$ \> Conditional equation \> [$-$] \\[0.5ex]
%  $e$ \> Willans coefficient \> [$-$] \\[0.5ex]
%  $F,G$ \> Parts of the system equation \> [\unitfrac[]{K}{s}]
% \end{tabbing}

% \section*{Indices}
% \begin{tabbing}
%  \hspace*{1.6cm}  \= \kill
%  a \> Ambient \\[0.5ex]
%  air \> Air
% \end{tabbing}

% \section*{Acronyms and Abbreviations}
% \begin{tabbing}
%  \hspace*{1.6cm}  \= \kill
%  NEDC \> New European Driving Cycle \\[0.5ex]
%  ETH \> Eidgen\"{o}ssische Technische Hochschule
% \end{tabbing}

%  \cleardoublepage

%---------------------------------------------------------------------------
